Exploiting the Social Graph: Increasing Engagement in Collaborative Interactive Evolution

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Response to comments
Answering some questions
And extending some explanations. Please @mariosky check them out if
you can.

IEAs also share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
and the need to build trust between participants and application providers. 
For the provider there is also the difficulty of establishing 
the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}. In order to increase participation and 
engagement, and to tackle the issues mentioned above,  
IEAs must follow a human centered design, % reference and explain what
                                % you mean by that. It's an EC
                                % conference - JJ
                                % I explain more later in other section, with references, maybe move that to here? -Mario
                                % If not I will add a few words later, such a large topic -Mario
                                % even with a Journal: http://hcis-journal.springeropen.com/about
giving extensive attention 
to volunteer users, not only because their
explicit evaluation is essential, but also because the context of the 
interaction affects the system as a whole. Users interact not only with the graphical
interface, and each phenotype, there is also the interaction
with other users, their previous evaluations and their own stream of activities
and past experiences. Thus users in the method presented in this paper are defined as dynamic entities 
interacting with an evolutionary computation having the fundamental purpose 
of evaluating phenotypes according to their preferences,
both explicitly or implicitly.

In this work we propose giving the same
importance to users and their interactions as the population of 
phenotypes have in a traditional EA, even to the extent of using this
knowledge as an integral part of the evolutionary algorithm. % how?
                                % Explain a bit more
                                % This is the idea: -M
Fitness assignment could depend on the actions of a group of users, and depend on how
they tag, share, rate, or even delete the phenotype. The selection of parents could
also depend on the previous actions, leveraging information such as the fac that  they have the same tag, or shared by
similar users, or whether a leader of opinion liked both or if they are stored in the same collection.
That knowledge could also be applied in choosing which phenotypes whould be presented to a
user, in order to give them those they will find more interesting or useful.

We argue also
%that a graph model is a viable representation for this purpose,  % define better the purpose. Not clear. "giving the same importance to users and their interactions"? - JJ
% I was thinking to record all the activities and relationships
that a graph model is a viable representation for the above relationships and
interactions and can then be used to model and use, within the
evolutionary algorithm, actions performed by every user. Graph based
models are currently used in social networks to keep track  
of user interactions with media objects, places, and other users. Users of
social networks (for instance the Facebook Graph) are accustomed to express these 
complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
Other example is the W3C Activity Streams 2.0 specification used for representing activities 
common in social web applications \cite{json:streams}..

%presenting a natural way of mapping the relationships found in the system. 
%Most IEAs have % have is not very precise and I'm a bit lost
               % here. What do you mean? is this constrained to IEAs
               % or EAs? At least the first is not exclusive, right? - JJ
%the following many-to-many relations:
When considering the type of relationships found in a genetic algorithm 
we can think of the Parent-Child relationship between a pair of parent solutions and
the child solution they produce. In IEAs there could be more entities involved, with 
even more types of interaction:

@@

  In IEAs these
  relationships are used to assign a single or a collection of
  fitness values to phenotypes \cite{garcia2013evospace}.   % I am
                                % afraid that individual is too
                                % similar to user, and you might get
                                % lost here. Maybe talk about
                                % "genotypes" or "phenotypes"? - JJ
                                % changed to phenotypes 
                                % do you think an additional explanation is needed? -Mario

@@

A case study is presented as a proof of concept, providing both
conceptual and implementation details of a graph data model applied in the 
implementation of an IEA. The experiment was implemented using
the EvoSpace-Interactive framework, and consisted of a web application
for evolving artistic drawings. The graph database is then used to
measure the degree of relationship between users and their participation.

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be shared with others. Collections can contain
phenotypes and a single phenotype could be shared by many
collections. % This is why they are not attributes - Mario
% You can simply have an attribute that is a list of the collections
% it belongs to - JJ
\end{itemize}

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.
%this is a bipartite graph them. - JJ
% Do you mean like a figure?  - Mario
% No, it's a kind of graph - JJ

\item {\bf Knows} The relation connects two users that know each
other. 
% This relation is "unique" meaning that it cannot be repeated
% between a pair of nodes.  % Not too clear what you mean by that - JJ
% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
 % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
 % There is even cases where knows relationship only goes in one direction, this could be because
 % a user does not publishes their friends or one only follows another but are not friends of each other.  
% Deleting the restriction, I think is not important and can be seen
% as a graph.
% It's a directed, bipartite graph then - JJ

>>>>

IEAs also share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
and the need to build trust between participants and application providers. 
For the provider there is also the difficulty of establishing 
the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}. In order to increase participation and 
engagement, and to tackle the issues mentioned above,  
IEAs must follow a human centered design, % reference and explain what
                                % you mean by that. It's an EC
                                % conference - JJ
                                % I explain more later in other section, with references, maybe move that to here? -Mario
                                % If not I will add a few words later, such a large topic -Mario
                                % even with a Journal:
                                % http://hcis-journal.springeropen.com/about
% a reference here would do no harm - JJ
giving extensive attention 
to volunteer users, not only because their
explicit evaluation is essential, but also because the context of the 
interaction affects the system as a whole. Users interact not only with the graphical
interface, and each phenotype, there is also the interaction
with other users, their previous evaluations and their own stream of activities
and past experiences. Thus users in the method presented in this paper are defined as dynamic entities 
interacting with an evolutionary computation having the fundamental purpose 
of evaluating phenotypes according to their preferences,
both explicitly or implicitly.

In this work we propose giving the same
importance to users and their interactions as the population of 
phenotypes have in a traditional EA, even to the extent of using this
knowledge as an integral part of the evolutionary algorithm.
Instead of letting the user assign fitness directly via a rating
or selection system, fitness assignment could depend on the actions of
a group of users, possibly connected in a social network, and depend on how 
they tag, share, rate, or even delete the phenotype. The selection of parents could
also depend on the previous actions, leveraging information such as the fac that  they have the same tag, or shared by
similar users, or whether a leader of opinion liked both or if they are stored in the same collection.
That knowledge could also be applied in choosing which phenotypes whould be presented to a
user, in order to give them those they will find more interesting or useful.

We argue also
that a graph model is a viable representation for the above relationships and
interactions and can then be used to model and use, within the
evolutionary algorithm, actions performed by every user, which, as
mentioned above, will be used to assign fitness to solutions. Graph based
models are currently used in social networks to keep track  
of user interactions with media objects, places, and other users. Users of
social networks (for instance the Facebook Graph) are accustomed to express these 
complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
Other example is the W3C Activity Streams 2.0 specification used for representing activities 
common in social web applications \cite{json:streams}.

When considering the type of relationships found in a genetic algorithm 
we can think of the Parent-Child relationship between a pair of parent solutions and
the child solution they produce. In IEAs there could be more entities involved, with 
even more types of interaction:

@@

  In IEAs these
  relationships are used to assign a single or a collection of
  fitness values to phenotypes \cite{garcia2013evospace}.   % I am
                                % afraid that individual is too
                                % similar to user, and you might get
                                % lost here. Maybe talk about
                                % "genotypes" or "phenotypes"? - JJ
                                % changed to phenotypes 
                                % do you think an additional
                                % explanation is needed? -Mario
  % Maybe explain that phenotypes are the expression of solutions
  % stored in the population, but do that above - JJ

@@

A case study is presented as a proof of concept, providing both
conceptual and implementation details of a graph data model applied in the 
implementation of an IEA. The experiment was implemented using
the EvoSpace-Interactive framework \cite{garcia2013evospace}, and consisted of a web application
for evolving artistic drawings. The graph database is then used to
measure the degree of relationship between users and their participation.

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be shared with others. Collections can contain
phenotypes and a single phenotype could be shared by many
collections. % This is why they are not attributes - Mario
% You can simply have an attribute that is a list of the collections
% it belongs to - JJ
% Right, but:
%  1) The relationship could have additional atributes, (when it was stored, order, etc.)
%  2) Then it could be difficult to query with Cypher (I think) - Mario
\end{itemize}

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.
%this is a bipartite graph them. - JJ
% Do you mean like a figure?  - Mario
% No, it's a kind of graph - JJ
% I read graph them :)  but is:  graph then , now I get it.

\item {\bf Knows} The relation connects two users that know each
other. 
% This relation is "unique" meaning that it cannot be repeated
% between a pair of nodes.  % Not too clear what you mean by that - JJ
% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
 % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
 % There is even cases where knows relationship only goes in one direction, this could be because
 % a user does not publishes their friends or one only follows another but are not friends of each other.  
% Deleting the restriction, I think is not important and can be seen
% as a graph.
% It's a directed, bipartite graph then - JJ
% Ok, I will mention the types of graphs. -Mario
